Standardized crypto-loans on\\ the Cardano blockchain

### 2021-02-03 papers/actus2021/article.tex
bib, formatting, overview, edited most of section 2.

\documentclass[runningheads]{llncs}
\usepackage[latin9]{inputenc}
\usepackage{url}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[unicode=true,
 bookmarks=false,
 breaklinks=false,pdfborder={0 0 1},backref=section,colorlinks=false]
 {hyperref}

@@

@@

%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Crypto-loans are unique and innovative financial instruments that
allow trustless P2P lending, which could provide a safe and convenient
source of liquidity for cryptocurrency holders. In this article, we
explore a smart-contract framework for building standardized crypto-loans
with Marlowe DSL and Algorithmic Contract Types Unified Standards
at its core.

@@

@@

List of contributions:
\begin{itemize}
\item \textquotedblleft Marlowe Labs\textquotedblright{} editor on Playground
(Dmytro) 
\item R Shiny cashflow visualiser for ACTUS (Dmytro) 
\item ACTUS Utility functions (Dmytro, Quanterall) 
\item executable ACTUS spec for PAM (Dmytro, Alex) 
\item executable ACTUS spec for LAM (Wyo Hackathon) 
\item Library for generating ACTUS contracts (Dmytro) 
\item ACTUS-like examples on Playground (Dmytro, Alex, Pablo) 
\item Marlowe DSL enhancements for Actus: Cond, Mul (Dmytro) (draft) 
\item ACTUS spec generator for Agda (Dmytro) 
\item Auto-refund static analysis (Pablo) (partial: only QC) 
\item Verification with QC + SMT (Dmytro) 
\end{itemize}

\section{Background}

@@

Cryptocurrency-backed loans must have collaterals as there is no trust
between party and counterparty. While loan is usually settled in a
stable-coin currency (e.g. USDT/USDC), the typical collateral is denominated
in a cryptocurrency (e.g. BTC). The purpose of this loan is to give
the borrower access to the fiat value of their crypto-funds without
actually selling them for fiat. Borrower pays interest In exchange
for provided liquidity.

Every loan has a positive net payoff (return minus investment) that
is either rendered as a one-time payment (see ZCB) or as a scheduled
payment of the interest. The rate of interest could be fixed throughout
the lifetime of a contract - for example zero-risk bonds have an fixed
interest proportional to inflation rate.

However, in a generic case - the interest rate is variable and depends
on an external factor agreed in advance, while the interest is periodically
updated by observing the state of that factor. Such loans usually
represent an investment in a particular venture or the industry. As
a somewhat fictional example, one could imagine a cryptocurrency miner
who decided to scale their crypto-farm: the loan (in USD) with variable
interest that directly depends on cryptocurrency prices would be more
attractive for a miner because it would directly correlate with miner's
profits. For example, if the price of the cryptocurrency goes down
in a particular month - miners would have to pay lower interest -
one always pays ``fixed share'' of the profits. In a more traditional
setup - the canonical example would be a power plant taking a loan
with interest depending on electricity prices. In both cases, either
cryptocurrency or electricity prices become a driver for the interest
rate.

Although, one can't just take the bare price of the asset and turn
it into a rate. In order to make ``units of measurement'' compatible
- adjustments should be made. Fluctuations of the interest rate driver
are embedded as:

\noindent 
\begin{equation}
\Delta_{r}=capfloor(driver*multiplier+spread-interestRate_{t-1})
\end{equation}

\noindent 
\begin{equation}
interestRate_{t+1}=capfloor(interestRate_{t-1}+\Delta_{r}),
\end{equation}

Where capfloor is a function that limits the range of fluctuation
(thus limiting lender's risk exposure):

\noindent 
\begin{equation}
capfloor(x)=max(min(x,floor),cap)
\end{equation}

Here, the spread parameter loosely represents the difference between
the average prime rate that the lender expects (benchmark yield) and
the rate imposed by the driver - the higher the spread, the higher
the resulting interest rate is. Multiplier effectively rescales the
interest rate plot in order to represent the difference between units
of measurement: how much percentage points you would get for a usd-to-kilowatt
and so on.

@@

Counterparty might decide to reinvest profit received as interest
from the same loan. In the simplest case, this renders as compound
interest. This can be modeled through interest accrual and capitalisation,
for instance contracts from the ACTUS specification accrue interest
between interest payments and can transfer interest to a notional
during interest capitalisation event (IPCL).

@@

\section{Modeling financial contracts}

\subsection{Marlowe }

@@

\section{Building an executable specification of ACTUS}

@@

\section{Generating Marlowe contracts from standardised ACTUS contract terms}

@@

\noindent 
\begin{equation}
chainlink(t)=receiveData(t)\circ calculatePayoff(t)\circ processPayoff(t)
\end{equation}

@@

where $receiveData$ - is asking an oracle for Marlowe Choice if needed,
$calculatePayoff$ - calculates payoff formula. For fixed-rate contracts
this is optimised into a pre-calculated constant. $processPayoff$
- awaits Deposit of payoff amount from a party: if deposit is transferred
it redirects the funds to a counterparty - otherwise it transfers
the collateral to a counterparty and closes the contract. The chain
is generated from the fixed (known in advance) schedule of events
using $SCHED()$function from ACTUS.

@@

\begin{verbatim}
if condition 
	then 
		perform_something1()
		continue() 
	else 
		perform_something2 
		continue()
\end{verbatim}

would inline contents of $continue()$ twice and given that ACTUS
contracts are essentially generated using continuation as an accumulator
- this would lead to exponential explosion of the size of any actus
contract that has conditionals in their state transition logic.

@@

We addressed this issue by introducing ``Cond'' construct in order
to represent ``if'' expressions. So instead of using the ``If''
contract to decide the value of some variable - we use conditional
assignment instead as ``Cond'' is essentially just a pure function
that returns value depending on condition in contrast to the ``If''
contract that performs action instead of returning value. The name
is borrowed from a similar function in Lisp.

\subsection{Limitations due to termination }

Marlowe doesn't allow perpetual contracts even if their recursion
is productive (example: perpetual swap) - thus we can't support certain
contract types from ACTUS specification, namely the ones that don't
have a defined maturity date (like UMP). In a long-term perspective
- there is a possible workaround: contracts with no maturity date
could be represented as actors with finite amounts of state transitions.
We prototyped this approach in a separate branch, however it does
seem more prone to errors comparing to rendering predefined schedules
and more importantly it greatly affects static analysis as the amount
of ``reduction steps'' grows from $N_{scheduled}=count(event)$
to $N_{stateTransitions}=\frac{max(date(event))-min(date(event))}{precision}$.

@@

\section{Tokenization}

@@

\section{Verification and testing}

@@

\section{Comparison with existing smart-contracts approaches }

@@

\section{Conclusion}

>>>>

\documentclass[runningheads]{llncs}
\usepackage[latin9]{inputenc}
\usepackage{url}[obeyspaces]
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[unicode=true,
 bookmarks=false,
 breaklinks=false,pdfborder={0 0 1},backref=section,colorlinks=false]
 {hyperref}

@@

% SJT use this to get nicer formatting of textual identifiers in math mode
\newcommand{\ident}[1]{\mbox{\emph{#1}}}

@@

%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Crypto-loans are %unique and 
innovative financial instruments that
allow trustless peer-to-peer %P2P 
lending, and potentially providing a safe and convenient
source of liquidity for cryptocurrency holders. In this paper we
explore a smart contract framework for building standardized crypto-loans
using the Marlowe domain-specific language and with the ACTUS standard at its core.

@@

\textbf{DONE:}
\begin{enumerate}
\item
Added overview at the end of the intro.
\item
Tidying up identifiers in math equations using the command \verb+\ident{...}+
\item
Biblio sorted out ordering (added keys). shortened author and editor lists
\end{enumerate}

\textbf{TO DO:}
\begin{enumerate}
\item
Algorithmic Contract Types Unified Standards (ACTUS) -- for first use. Also reference.
\item
Do we include Alex as co-author?
\item
What about R (shiny)?
\end{enumerate}

@@

\bigskip\noindent
List of contributions:
\begin{itemize}
\item \textquotedblleft Marlowe Labs\textquotedblright{} editor on Playground
(Dmytro) 
\item R Shiny cashflow visualiser for ACTUS (Dmytro) 
\item ACTUS Utility functions (Dmytro, Quanterall) 
\item executable ACTUS spec for PAM (Dmytro, Alex) 
\item executable ACTUS spec for LAM (Wyo Hackathon) 
\item Library for generating ACTUS contracts (Dmytro) 
\item ACTUS-like examples on Playground (Dmytro, Alex, Pablo) 
\item Marlowe DSL enhancements for Actus: Cond, Mul (Dmytro) (draft) 
\item ACTUS spec generator for Agda (Dmytro) 
\item Auto-refund static analysis (Pablo) (partial: only QC) 
\item Verification with QC + SMT (Dmytro) 
\end{itemize}

% SJT: added this overview of the paper
% it is traditional to end the introduction like this,
% 'signposting' what is in the rest of the paper.
Section \ref{background} of the paper covers the relevant financial background, and Section \ref{modelling} introduces Marlowe as well as the ACTUS financial standard. Section \ref{executable} builds an executable specification of ACTUS in Haskell, and this is used in Section \ref{generation} to generate  the Marlowe code for an ACTUS contract from the terms of the contract. Section \ref{tokenization} explains how tokens are used to represent ownership of roles in a running contract, and Section \ref{verification} presents how we provide assurance of the contract correctness. Section \ref{related} examines related work and Section \ref{conclusion} concludes.

\section{Financial contracts}
\label{background}

@@

Cryptocurrency-backed loans must have \emph{collateral} when there is no trust
between party and counterparty. While a loan is usually settled in a
stable-coin currency (e.g. USDT/USDC), collateral is typically denominated
in a cryptocurrency (e.g. BTC). The purpose of such a loan is to give
the borrower access to the fiat value of their crypto-funds without
actually selling them for fiat. The borrower pays interest in exchange
for gaining liquidity.

Every loan has a positive net payoff (return minus investment) that
is either rendered as a one-time payment -- often called a \emph{zero-coupon bond} (ZCB) -- or by scheduling
payment of the interest. The rate of interest could be fixed throughout
the lifetime of a contract: for example, zero-risk bonds have a fixed
interest proportional to the inflation rate.
However, in the generic case the interest rate is variable and depends
on an external factor agreed in advance, and the rate is periodically
updated by observing the state of that factor.

Such loans often represent an investment in a particular venture or industry. As
a somewhat fictional example, one could imagine a cryptocurrency miner
who decided to scale their crypto-farm: a loan (in USD) with variable
interest that directly depends on cryptocurrency prices would be more
attractive for a miner because it would directly correlate with miner's
profits. For example, if the price of the cryptocurrency goes down
in a particular month the borrower would have to pay lower interest, and so would always pay a fixed share of the profits. In a more traditional
setup the interest rate could depend on prices of other commodities: the canonical example would be a power plant taking a loan
with interest depending on electricity prices.

In both cases, the prices of
cryptocurrency or electricity become a driver for the interest
rate.
However, one cannot simply take the bare price of the asset and turn
it into a rate. In order to make \emph{units of measurement} compatible
%SJT compatible with what??
adjustments should be made. Fluctuations of the interest rate driver
are embedded thus:

\noindent 
\begin{equation}
\Delta_{r}=\ident{capfloor}(\ident{driver}*\ident{multiplier}+\ident{spread}-\ident{interestRate}_{t-1})
\end{equation}

\noindent 
\begin{equation}
\ident{interestRate}_{t+1}=\ident{capfloor}(\ident{interestRate}_{t-1}+\Delta_{r}),
\end{equation}

\noindent
where \emph{capfloor} is a function that limits the range of fluctuation,  
and so limiting the lender's exposure to risk:

\noindent 
\begin{equation}
\ident{capfloor}(x)=\ident{max(min}(x,\ident{floor}),\ident{cap})
\end{equation}

\noindent
The \emph{spread} parameter here loosely represents the difference between
the average prime 
% SJT explain "prime" here.
rate that the lender expects (the benchmark yield) and
the rate imposed by the driver - the higher the spread, the higher
the resulting interest rate. The multiplier %effectively 
rescales the interest rate value 
% SJT used value here rather than plot: is that OK?
in order to represent the changes to be made converting between different units
of measurement: how many percentage points you would get for a usd-to-kilowatt conversion
and so on.

@@

Counterparty might decide to reinvest profit received as interest
from the same loan. In the simplest case, this renders as compound
interest. This can be modelled through interest accrual and capitalisation,
for instance contracts from the ACTUS specification accrue interest
between interest payments and can transfer interest to a notional
during interest capitalisation event (IPCL).

@@

\section{Modelling financial contracts}
\label{modelling}

This section introduces the Marlowe DSL and the ACTUS standard for financial contracts.

\subsection{Marlowe}

@@

\section{Building an executable specification of ACTUS}
\label{executable}

@@

\section{Generating Marlowe contracts from ACTUS specs}%standardised ACTUS contract terms}
\label{generation}

@@

\noindent 
\begin{equation}
chainlink(t)=receiveData(t)\circ \mbox{\emph{calculatePayoff(t)}}\circ \mbox{\emph{processPayoff(t)}}
\end{equation}

@@

where $receiveData$ - is asking an oracle for Marlowe Choice if needed,
$\mbox{\emph{calculatePayoff}}$ - calculates payoff formula. For fixed-rate contracts
this is optimised into a pre-calculated constant. $\mbox{\emph{processPayoff}}$
- awaits Deposit of payoff amount from a party: if deposit is transferred
it redirects the funds to a counterparty - otherwise it transfers
the collateral to a counterparty and closes the contract. The chain
is generated from the fixed (known in advance) schedule of events
using the $SCHED()$ function from ACTUS.

@@

\begin{verbatim}
if condition 
	then 
		perform_something1()
		continue() 
	else 
		perform_something2 
		continue()
\end{verbatim}
\noindent
would inline contents of \texttt{continue()} twice and given that ACTUS
contracts are essentially generated using continuation as an accumulator
- this would lead to exponential explosion of the size of any ACTUS
contract that has conditionals in their state transition logic.

@@

We addressed this issue by introducing the \texttt{Cond} construct in order
to represent conditional expressions. So instead of using the \texttt{If}
contract to decide the value of some variable - we use conditional
assignment instead as \texttt{Cond} is essentially just a pure function
that returns value depending on condition in contrast to the \texttt{If}
contract that performs action instead of returning value. The name
is borrowed from a similar function in Lisp.

\subsection{Limitations due to termination}

Marlowe doesn't allow perpetual contracts even if their recursion
is productive (example: perpetual swap) - thus we can't support certain
contract types from ACTUS specification, namely the ones that don't
have a defined maturity date (like UMP). In a long-term perspective
- there is a possible workaround: contracts with no maturity date
could be represented as actors with finite amounts of state transitions.
We prototyped this approach in a separate branch, however it does
seem more prone to errors comparing to rendering predefined schedules
and more importantly it greatly affects static analysis as the amount
of ``reduction steps'' grows from 
\noindent 
\begin{equation}
N_{scheduled}=count(event)
\end{equation}
to 
\noindent 
\begin{equation}
N_{stateTransitions}=\frac{max(date(event))-min(date(event))}{precision} 
\end{equation}

@@

\section{Tokenization}
\label{tokenization}

@@

\section{Verification and testing}
\label{verification}

@@

\section{Related work}\label{related}

% SJT changed title from {Comparison with existing smart-contracts approaches }
% it's traditional to call this "relatyed work".

@@

\section{Conclusion}
\label{conclusion}
